**Title:** Analyzing long-term water quality of lakes in Rhode Island and the Northeastern United States with an anomaly approach

### 2021-04-15 manuscript/ecosphere_production/manuscript.Rmd
Updated table references

In general, there was little evidence to suggest broad regional changes in chlorophyll *a* as it showed a very weak positive trend, slight statistical support and above average years spread evenly throughout the time period (slope = `r signif(lagos_chla_rl$slope,2)`, p = `r signif(lagos_chla_rl$p.value, 2)`, `r fig_num(name = "chla_trend", display = "cite")`B., `r tab_num(name = "trend_summ", display = "cite")`).  Total nitrogen showed a slight decreasing trend with weak statistical support and slope less than the 3.5 μg/L that would result in a trophic state change over a century, thus there is little support for a meaningful trend in TN at the regional scale (slope = `r signif(lagos_tn_rl$slope, 2)`, p = `r signif(lagos_tn_rl$p.value, 2)`, `r fig_num(name = "tn_trend", display = "cite")`B., `r tab_num(name = "trend_summ", display = "cite")`).  Furthermore, The last two years for which LAGOSNE had data on total nitrogen were from a relatively smaller number of sites (`r fig_num("year_n", display = "cite")`) and those values may not be representative.  Total phosphorus showed a very small increasing trend and there is statistical support for that trend.  The slope however suggests little support for an ecological meaningful change in total phosphorus (slope = `r signif(lagos_tp_rl$slope, 2)`, p = `r signif(lagos_tp_rl$p.value,2)`, `r fig_num(name = "tp_trend", display = "cite")`B., `r tab_num(name = "", display = "cite")`). Lastly, the TN:TP ratio showed little change (slope = `r signif(lagos_np_rl$slope, 2)`, p = `r signif(lagos_np_rl$p.value, 2)`, `r fig_num(name = "np_trend", display = "cite")`B., `r tab_num(name = "
", display = "cite")`) as none of our criteria for a trend were met.  Although, there appears to be a non-linear trend that suggests increasing TN:TP early in the time period and decreasing TN:TP later, we feel we cannot effectively evaluate this though as total nitrogen measurements were rare in 2011 (`r fig_num("year_n", display = "cite")`) and sites that had both nitrogen and phosphorus were even rarer.  This low median anomaly should be interpreted with this caveat in mind.   Taken together we feel our results largely match the findings of Oliver et al. [-@Oliver_2017] that there is stasis in lake nutrients and chlorophyll within the LAGOSNE region.

>>>>

In general, there was little evidence to suggest broad regional changes in chlorophyll *a* as it showed a very weak positive trend, slight statistical support and above average years spread evenly throughout the time period (slope = `r signif(lagos_chla_rl$slope,2)`, p = `r signif(lagos_chla_rl$p.value, 2)`, `r fig_num(name = "chla_trend", display = "cite")`B., `r tab_num(name = "trend_summ", display = "cite")`).  Total nitrogen showed a slight decreasing trend with weak statistical support and slope less than the 3.5 μg/L that would result in a trophic state change over a century, thus there is little support for a meaningful trend in TN at the regional scale (slope = `r signif(lagos_tn_rl$slope, 2)`, p = `r signif(lagos_tn_rl$p.value, 2)`, `r fig_num(name = "tn_trend", display = "cite")`B., `r tab_num(name = "trend_summ", display = "cite")`).  Furthermore, The last two years for which LAGOSNE had data on total nitrogen were from a relatively smaller number of sites (`r fig_num("year_n", display = "cite")`) and those values may not be representative.  Total phosphorus showed a very small increasing trend and there is statistical support for that trend.  The slope however suggests little support for an ecological meaningful change in total phosphorus (slope = `r signif(lagos_tp_rl$slope, 2)`, p = `r signif(lagos_tp_rl$p.value,2)`, `r fig_num(name = "tp_trend", display = "cite")`B., `r tab_num(name = "trend_summ", display = "cite")`). Lastly, the TN:TP ratio showed little change (slope = `r signif(lagos_np_rl$slope, 2)`, p = `r signif(lagos_np_rl$p.value, 2)`, `r fig_num(name = "np_trend", display = "cite")`B., `r tab_num(name = "trend_summ", display = "cite")`) as none of our criteria for a trend were met.  Although, there appears to be a non-linear trend that suggests increasing TN:TP early in the time period and decreasing TN:TP later, we feel we cannot effectively evaluate this though as total nitrogen measurements were rare in 2011 (`r fig_num("year_n", display = "cite")`) and sites that had both nitrogen and phosphorus were even rarer.  This low median anomaly should be interpreted with this caveat in mind.   Taken together we feel our results largely match the findings of Oliver et al. [-@Oliver_2017] that there is stasis in lake nutrients and chlorophyll within the LAGOSNE region.
